### 2018-05-30 submission/manuscript.Rmd
update first discussion paragraph

In this study we show that the number of OTUs, error rate, and chimera prevalence depends on polymerase used [Figure 1-5]. These differences are important because many diversity metrics rely on the number of OTUs or other measures dependent on error rate and chimera prevalence as part of their metric calculations (e.g. richness). Additionally, the earlier detection of differences in total number of OTUs between polymerases in the mock versus fecal samples might indicate that high biomass samples may underestimate the biases present within low biomass samples. We observed that undetected chimeras that were not identified and removed using standard bioinformatic approaches cause many of these differences. This suggests that some of the specific diversity differences between studies can be attributed to differences in polyermase used. Based on these observations, metrics that measure within sample diversity like richness depend on polymerase but this may not be the case for metrics that assess between sample diversity.

>>>>

Our observations support that the number of OTUs, error rate, and chimera prevalence depends on polymerase used and that chimera prevalence is highly correlated with the number of OTUs [Figure 1-5]. Our observations also suggest that undetected chimeras, not removed using standard bioinformatic approaches, cause many of these differences. In addition to the reported polymerase dependent differences, the bacterial community also varied by the number of cycles used. These differences are important because many diversity metrics rely on calculations (e.g. richness) that are dependent on the number of OTUs. Since the number of OTUs are dependent on error rate and chimera prevalence the diversity metrics can vary simply by changing the polymerase used. Based on these observations, metrics that measure diversity using calculations such as richness depend on the polymerase, but this may not be the case for all diversity metrics.

### 2018-05-30 submission/manuscript.tex
update first discussion paragraph

In this study we show that the number of OTUs, error rate, and chimera
prevalence depends on polymerase used {[}Figure 1-5{]}. These
differences are important because many diversity metrics rely on the
number of OTUs or other measures dependent on error rate and chimera
prevalence as part of their metric calculations (e.g.~richness).
Additionally, the earlier detection of differences in total number of
OTUs between polymerases in the mock versus fecal samples might indicate
that high biomass samples may underestimate the biases present within
low biomass samples. We observed that undetected chimeras that were not
identified and removed using standard bioinformatic approaches cause
many of these differences. This suggests that some of the specific
diversity differences between studies can be attributed to differences
in polyermase used. Based on these observations, metrics that measure
within sample diversity like richness depend on polymerase but this may
not be the case for metrics that assess between sample diversity.

>>>>

Our observations support that the number of OTUs, error rate, and
chimera prevalence depends on polymerase used and that chimera
prevalence is highly correlated with the number of OTUs {[}Figure
1-5{]}. Our observations also suggest that undetected chimeras, not
removed using standard bioinformatic approaches, cause many of these
differences. In addition to the reported polymerase dependent
differences, the bacterial community also varied by the number of cycles
used. These differences are important because many diversity metrics
rely on calculations (e.g.~richness) that are dependent on the number of
OTUs. Since the number of OTUs are dependent on error rate and chimera
prevalence the diversity metrics can vary simply by changing the
polymerase used. Based on these observations, metrics that measure
diversity using calculations such as richness depend on the polymerase,
but this may not be the case for all diversity metrics.
